---
output: html_document
editor_options: 
  chunk_output_type: console
---
\newpage
\tableofcontents 
\newpage
\listoftables 
\newpage
\listoffigures 
\newpage

```{r setup, include=FALSE}
# Set your working directory
# Load packages
getwd()
library(tidyverse)
library(sf)
library(ggplot2)
library(dplyr)
library(mapview)
library(leaflet)
library("GGally")
library(readxl)
library(RColorBrewer)


# Set theme
mytheme <- theme_classic(base_size = 14) +
  theme(axis.text = element_text(color = "black"), 
        legend.position = "top")
theme_set(mytheme)

```

```{r}
# Read in census data
census_processed<- read.csv("Data/census_processed(2020-2024).csv", stringsAsFactors = T)

census_processed <- census_processed %>% 
  mutate(MEAN = rowMeans(across(X2020:X2024)))
substring(census_processed$State, 2, nchar(census_processed))
colnames(census_processed) <- c("State","2020","2021","2022","2023","2024")

#Federal Energy and Environment investment data by county 
Energy_investment<- read.csv("Data/Energy_investment_by_county.csv", stringsAsFactors = T)

#Federal Energy and Environment investment data by state 
Investment_state <- read.csv("Data/Raw/Federal energy and environment spending by STATE (2010-2025).csv", stringsAsFactors = T)

#Natural disaster data
disasters <- read.csv("Data/DisasterDeclarationsSummaries.csv", stringsAsFactors = T)

#spatial data
states <- st_read("Data/Spatial/cb_2018_us_state_20m.shp", stringsAsFactors = T)


census_2021 <- read.csv("Data/Census/ACSSPP1Y2021.S0201-2025-11-29T154845.csv", stringsAsFactors = T)
census_2022 <- read.csv("Data/Census/ACSSPP1Y2022.S0201-2025-11-29T154817.csv", stringsAsFactors = T)
census_2023 <- read.csv("Data/Census/ACSSPP1Y2023.S0201-2025-11-29T154759.csv", stringsAsFactors = T)
census_2024 <- read.csv("Data/Census/ACSSPP1Y2024.S0201-2025-11-29T154742.csv", stringsAsFactors = T)
```



```{r}
#Data Wrangling (for 2021)
##Clean the data

#census_2021.columns = census_2021.columns.str.replace(' ','')
census2021 <- as.data.frame(t(census_2021))
colnames(census2021) <-c(census2021[1,])

census2021clean <- census2021[-1,]
#census2021clean <- census2021[-c(seq(2,length(census2021),2)),]

colnames(census2021clean) <- str_trim(colnames(census2021clean))

rownames(census2021clean) <- str_sub(rownames(census2021clean), end =-29)
#rownames(census2021clean) <- gsub("."," ", rownames(census2021clean))

census2021clean <- census2021clean %>% 
  select("Total population", "Male", "Female", "Under 5 years", "5 to 17 years", "Median age (years)", "Median household income (dollars)","Gas", "Electricity","All other fuels","No fuel used") %>% 
  mutate(year = 2021, State = rownames(census2021clean))
  #mutate(State = sub("."," ", State)) # Remove the dots in State names
census2021clean$State <- gsub( "\\."," ",census2021clean$State)

census2021clean$`Total population` <- as.numeric(str_trim(census2021clean$`Total population`))
census2021clean$`Male` <- as.numeric(str_trim(census2021clean$`Male`))
census2021clean$Female <- as.numeric(str_trim(census2021clean$`Female`))
census2021clean$`Under 5 years` <- as.numeric(str_trim(census2021clean$`Under 5 years`))
census2021clean$`5 to 17 years` <- as.numeric(str_trim(census2021clean$`5 to 17 years`))
census2021clean$`Median age (years)` <- as.numeric(str_trim(census2021clean$`Median age (years)`))
census2021clean$`Median household income (dollars)` <- as.numeric(str_trim(census2021clean$`Median household income (dollars)`))

census2021clean$Gas <- as.numeric(str_trim(census2021clean$Gas))
census2021clean$Electricity <- as.numeric(str_trim(census2021clean$Electricity))
census2021clean$`All other fuels` <- as.numeric(str_trim(census2021clean$`All other fuels`))
census2021clean$`No fuel used` <- as.numeric(str_trim(census2021clean$`No fuel used`))


#add column for children ages 0-17 

census2021clean <- census2021clean%>%
  mutate(child=`Under 5 years` + `5 to 17 years`)


## Energy dataset (State Level)
Investment <- Investment_state %>%  #data for time series
  select(
    "TotalFedAction_EnergyEnv_2010","TotalFedAction_2010",
    "TotalFedAction_EnergyEnv_2011","TotalFedAction_2011",
    "TotalFedAction_EnergyEnv_2012","TotalFedAction_2012",
    "TotalFedAction_EnergyEnv_2013","TotalFedAction_2013",
    "TotalFedAction_EnergyEnv_2014","TotalFedAction_2014",
    "TotalFedAction_EnergyEnv_2015","TotalFedAction_2015",
    "TotalFedAction_EnergyEnv_2016","TotalFedAction_2016",
    "TotalFedAction_EnergyEnv_2017","TotalFedAction_2017",
    "TotalFedAction_EnergyEnv_2018","TotalFedAction_2018",
    "TotalFedAction_EnergyEnv_2019","TotalFedAction_2019",
    "TotalFedAction_EnergyEnv_2020","TotalFedAction_2020",
    "TotalFedAction_EnergyEnv_2021","TotalFedAction_2021",
    "TotalFedAction_EnergyEnv_2022","TotalFedAction_2022",
    "TotalFedAction_EnergyEnv_2023","TotalFedAction_2023",
    "TotalFedAction_EnergyEnv_2024","TotalFedAction_2024") %>% 
  mutate(across(contains("FedAction"),
                ~ as.numeric(str_replace_all(.x, "[\\$,]", "")))) 
  
  
Investment_state_2021 <- Investment_state %>% 
  select("STATEFP","stname", "stusps","TotalFedAction_EnergyEnv_2021","TotalFedAction_2021")
Investment_state_2021$TotalFedAction_EnergyEnv_2021 <- 
gsub("\\$ ", "", Investment_state_2021$TotalFedAction_EnergyEnv_2021)
Investment_state_2021$TotalFedAction_2021 <- 
gsub("\\$ ", "", Investment_state_2021$TotalFedAction_2021)
Investment_state_2021 <- Investment_state_2021 %>% 
  mutate(TotalFedAction_EnergyEnv_2021 =
           as.numeric(
             gsub(",", "", trimws(TotalFedAction_EnergyEnv_2021)))
         ) %>% 
  mutate(TotalFedAction_2021 =
           as.numeric(
             gsub(",", "", trimws(TotalFedAction_2021)))
         ) %>% 
  mutate(Ratio = TotalFedAction_EnergyEnv_2021/TotalFedAction_2021,
         Year = 2021)


#Disaster
states_names <- data.frame(abbrev = states$STUSPS,
                           State = states$NAME) 
#The disaster data set has only abbreviation for states, so full names need to be joined from other dataset.

#summarize disaster count by state
disaster2021 <- disasters %>% 
  filter(fyDeclared %in% c(2021)) %>% 
  arrange(state) %>% 
  group_by(state, fipsStateCode) %>% 
  summarise(Disasters = n())

disaster2021 <- left_join(disaster2021, states_names, by = c("state" = "abbrev"))

#Public awareness
Regulate <- read_excel("Data/Regulate_percentage(2021~2024).xlsx")
Regulate2021 <- Regulate %>% 
  select("GeoName","2021")


#Combining all the variables (Demographic, disaster and public awareness)
combined2021 <- left_join(census2021clean, Investment_state_2021, by = c("State" = "stname"))

combined2021 <- left_join(combined2021, disaster2021, by = c("State" = "State"))

combined2021 <- left_join(combined2021, Regulate2021, by = c("State" = "GeoName"))



write.csv(combined2021, file = "Data/Processed/combined2021.csv")
```

```{r}
#Data Wrangling (for 2022)
##Clean the data

census2022 <- as.data.frame(t(census_2022))
colnames(census2022) <- c(census2022[1,])

census2022clean <- census2022[-1,]

colnames(census2022clean) <- str_trim(colnames(census2022clean))

rownames(census2022clean) <- str_sub(rownames(census2022clean), end = -29)

census2022clean <- census2022clean %>% 
  select("Total population", "Male", "Female", "Under 5 years", "5 to 17 years",
         "Median age (years)", "Median household income (dollars)",
         "Gas", "Electricity", "All other fuels", "No fuel used") %>% 
  mutate(year = 2022, State = rownames(census2022clean))

census2022clean$State <- gsub("\\.", " ", census2022clean$State)

# Numeric conversions
cols_to_num <- c("Total population","Male","Female","Under 5 years","5 to 17 years",
                 "Median age (years)","Median household income (dollars)",
                 "Gas","Electricity","All other fuels","No fuel used")

census2022clean[cols_to_num] <- lapply(census2022clean[cols_to_num],
                                       function(x) as.numeric(str_trim(x)))

# Child column
census2022clean <- census2022clean %>%
  mutate(child = `Under 5 years` + `5 to 17 years`)


## Energy dataset (State Level)
Investment <- Investment_state %>% 
  select(
    "TotalFedAction_EnergyEnv_2010","TotalFedAction_2010",
    "TotalFedAction_EnergyEnv_2011","TotalFedAction_2011",
    "TotalFedAction_EnergyEnv_2012","TotalFedAction_2012",
    "TotalFedAction_EnergyEnv_2013","TotalFedAction_2013",
    "TotalFedAction_EnergyEnv_2014","TotalFedAction_2014",
    "TotalFedAction_EnergyEnv_2015","TotalFedAction_2015",
    "TotalFedAction_EnergyEnv_2016","TotalFedAction_2016",
    "TotalFedAction_EnergyEnv_2017","TotalFedAction_2017",
    "TotalFedAction_EnergyEnv_2018","TotalFedAction_2018",
    "TotalFedAction_EnergyEnv_2019","TotalFedAction_2019",
    "TotalFedAction_EnergyEnv_2020","TotalFedAction_2020",
    "TotalFedAction_EnergyEnv_2021","TotalFedAction_2021",
    "TotalFedAction_EnergyEnv_2022","TotalFedAction_2022",
    "TotalFedAction_EnergyEnv_2023","TotalFedAction_2023",
    "TotalFedAction_EnergyEnv_2024","TotalFedAction_2024") %>% 
  mutate(across(contains("FedAction"),
                ~ as.numeric(str_replace_all(.x, "[\\$,]", "")))) 


Investment_state_2022 <- Investment_state %>% 
  select("STATEFP","stname","stusps",
         "TotalFedAction_EnergyEnv_2022","TotalFedAction_2022")

Investment_state_2022$TotalFedAction_EnergyEnv_2022 <- 
  gsub("\\$ ", "", Investment_state_2022$TotalFedAction_EnergyEnv_2022)

Investment_state_2022$TotalFedAction_2022 <- 
  gsub("\\$ ", "", Investment_state_2022$TotalFedAction_2022)

Investment_state_2022 <- Investment_state_2022 %>% 
  mutate(TotalFedAction_EnergyEnv_2022 =
           as.numeric(gsub(",", "", trimws(TotalFedAction_EnergyEnv_2022)))) %>% 
  mutate(TotalFedAction_2022 =
           as.numeric(gsub(",", "", trimws(TotalFedAction_2022)))) %>% 
  mutate(Ratio = TotalFedAction_EnergyEnv_2022 / TotalFedAction_2022,
         Year = 2022)


#Disaster
states_names <- data.frame(abbrev = states$STUSPS,
                           State = states$NAME)

disaster2022 <- disasters %>% 
  filter(fyDeclared %in% c(2022)) %>% 
  arrange(state) %>% 
  group_by(state, fipsStateCode) %>% 
  summarise(Disasters = n())

disaster2022 <- left_join(disaster2022, states_names, by = c("state" = "abbrev"))

#Public awareness
Regulate <- read_excel("Data/Regulate_percentage(2021~2024).xlsx")
Regulate2022 <- Regulate %>% 
  select("GeoName","2022")


#Combined dataset
combined2022 <- left_join(census2022clean, Investment_state_2022,
                          by = c("State" = "stname"))

combined2022 <- left_join(combined2022, disaster2022,
                          by = c("State" = "State"))

combined2022 <- left_join(combined2022, Regulate2022,
                          by = c("State" = "GeoName"))

write.csv(combined2022, file = "Data/Processed/combined2022.csv")

```

```{r}
# Data Wrangling (for 2023)
## Clean the data

census2023 <- as.data.frame(t(census_2023))
colnames(census2023) <- c(census2023[1,])
census2023clean <- census2023[-1,]

colnames(census2023clean) <- str_trim(colnames(census2023clean))
rownames(census2023clean) <- str_sub(rownames(census2023clean), end = -29)

census2023clean <- census2023clean %>% 
  select("Total population", "Male", "Female", "Under 5 years", "5 to 17 years",
         "Median age (years)", "Median household income (dollars)",
         "Gas", "Electricity", "All other fuels", "No fuel used") %>% 
  mutate(year = 2023, State = rownames(census2023clean))

census2023clean$State <- gsub("\\.", " ", census2023clean$State)

# Numeric conversions
cols_to_num <- c("Total population","Male","Female","Under 5 years","5 to 17 years",
                 "Median age (years)","Median household income (dollars)",
                 "Gas","Electricity","All other fuels","No fuel used")

census2023clean[cols_to_num] <- lapply(census2023clean[cols_to_num],
                                       function(x) as.numeric(str_trim(x)))

# Child column
census2023clean <- census2023clean %>%
  mutate(child = `Under 5 years` + `5 to 17 years`)


## Energy dataset (State Level)
Investment_state_2023 <- Investment_state %>% 
  select("STATEFP","stname","stusps",
         "TotalFedAction_EnergyEnv_2023","TotalFedAction_2023")

Investment_state_2023$TotalFedAction_EnergyEnv_2023 <- 
  gsub("\\$ ", "", Investment_state_2023$TotalFedAction_EnergyEnv_2023)
Investment_state_2023$TotalFedAction_2023 <- 
  gsub("\\$ ", "", Investment_state_2023$TotalFedAction_2023)

Investment_state_2023 <- Investment_state_2023 %>% 
  mutate(TotalFedAction_EnergyEnv_2023 =
           as.numeric(gsub(",", "", trimws(TotalFedAction_EnergyEnv_2023)))) %>% 
  mutate(TotalFedAction_2023 =
           as.numeric(gsub(",", "", trimws(TotalFedAction_2023)))) %>% 
  mutate(Ratio = TotalFedAction_EnergyEnv_2023 / TotalFedAction_2023,
         Year = 2023)


# Disaster
states_names <- data.frame(abbrev = states$STUSPS,
                           State = states$NAME)

disaster2023 <- disasters %>% 
  filter(fyDeclared %in% c(2023)) %>% 
  arrange(state) %>% 
  group_by(state, fipsStateCode) %>% 
  summarise(Disasters = n())

disaster2023 <- left_join(disaster2023, states_names, by = c("state" = "abbrev"))

# Public awareness
Regulate <- read_excel("Data/Regulate_percentage(2021~2024).xlsx")
Regulate2023 <- Regulate %>% 
  select("GeoName","2023")


# Combined dataset
combined2023 <- left_join(census2023clean, Investment_state_2023,
                          by = c("State" = "stname"))

combined2023 <- left_join(combined2023, disaster2023,
                          by = c("State" = "State"))

combined2023 <- left_join(combined2023, Regulate2023,
                          by = c("State" = "GeoName"))

write.csv(combined2023, file = "Data/Processed/combined2023.csv")

```

```{r}
# Data Wrangling (for 2024)
## Clean the data

census2024 <- as.data.frame(t(census_2024))
colnames(census2024) <- c(census2024[1,])
census2024clean <- census2024[-1,]

colnames(census2024clean) <- str_trim(colnames(census2024clean))
rownames(census2024clean) <- str_sub(rownames(census2024clean), end = -29)

census2024clean <- census2024clean %>% 
  select("Total population", "Male", "Female", "Under 5 years", "5 to 17 years",
         "Median age (years)", "Median household income (dollars)",
         "Gas", "Electricity", "All other fuels", "No fuel used") %>% 
  mutate(year = 2024, State = rownames(census2024clean))

census2024clean$State <- gsub("\\.", " ", census2024clean$State)

# Numeric conversions
cols_to_num <- c("Total population","Male","Female","Under 5 years","5 to 17 years",
                 "Median age (years)","Median household income (dollars)",
                 "Gas","Electricity","All other fuels","No fuel used")

census2024clean[cols_to_num] <- lapply(census2024clean[cols_to_num],
                                       function(x) as.numeric(str_trim(x)))

# Child column
census2024clean <- census2024clean %>%
  mutate(child = `Under 5 years` + `5 to 17 years`)


## Energy dataset (State Level)
Investment_state_2024 <- Investment_state %>% 
  select("STATEFP","stname","stusps",
         "TotalFedAction_EnergyEnv_2024","TotalFedAction_2024")

Investment_state_2024$TotalFedAction_EnergyEnv_2024 <- 
  gsub("\\$ ", "", Investment_state_2024$TotalFedAction_EnergyEnv_2024)
Investment_state_2024$TotalFedAction_2024 <- 
  gsub("\\$ ", "", Investment_state_2024$TotalFedAction_2024)

Investment_state_2024 <- Investment_state_2024 %>% 
  mutate(TotalFedAction_EnergyEnv_2024 =
           as.numeric(gsub(",", "", trimws(TotalFedAction_EnergyEnv_2024)))) %>% 
  mutate(TotalFedAction_2024 =
           as.numeric(gsub(",", "", trimws(TotalFedAction_2024)))) %>% 
  mutate(Ratio = TotalFedAction_EnergyEnv_2024 / TotalFedAction_2024,
         Year = 2024)


# Disaster
states_names <- data.frame(abbrev = states$STUSPS,
                           State = states$NAME)

disaster2024 <- disasters %>% 
  filter(fyDeclared %in% c(2024)) %>% 
  arrange(state) %>% 
  group_by(state, fipsStateCode) %>% 
  summarise(Disasters = n())

disaster2024 <- left_join(disaster2024, states_names, by = c("state" = "abbrev"))

# Public awareness
Regulate <- read_excel("Data/Regulate_percentage(2021~2024).xlsx")
Regulate2024 <- Regulate %>% 
  select("GeoName","2024")


# Combined dataset
combined2024 <- left_join(census2024clean, Investment_state_2024,
                          by = c("State" = "stname"))

combined2024 <- left_join(combined2024, disaster2024,
                          by = c("State" = "State"))

combined2024 <- left_join(combined2024, Regulate2024,
                          by = c("State" = "GeoName"))

write.csv(combined2024, file = "Data/Processed/combined2024.csv")

```

```{r}
#Combine the four years
#2021
df_2021 <- combined2021 %>%
  rename(
    Total_EE_Investment = TotalFedAction_EnergyEnv_2021, 
    Political_Score = `2021`                            
  ) %>%
  mutate(Year = 2021) %>% 
  select(State, Year, Total_EE_Investment, Political_Score, Disasters, 
         `Total population`, Female, child, `Median age (years)`, 
         `Median household income (dollars)`, Gas, Electricity) 
      

# 2022
df_2022 <- combined2022 %>%
  rename(
    Total_EE_Investment = TotalFedAction_EnergyEnv_2022,
    Political_Score = `2022`
  ) %>%
  mutate(Year = 2022) %>% 
  select(State, Year, Total_EE_Investment, Political_Score, Disasters, 
         `Total population`, Female, child, `Median age (years)`, 
         `Median household income (dollars)`, Gas, Electricity)

# 2023
df_2023 <- combined2023 %>%
  rename(
    Total_EE_Investment = TotalFedAction_EnergyEnv_2023,
    Political_Score = `2023`
  ) %>%
  mutate(Year = 2023) %>% 
  select(State, Year, Total_EE_Investment, Political_Score, Disasters, 
         `Total population`, Female, child, `Median age (years)`, 
         `Median household income (dollars)`, Gas, Electricity)

# 2024
df_2024 <- combined2024 %>%
  rename(
    Total_EE_Investment = TotalFedAction_EnergyEnv_2024,
    Political_Score = `2024`
  ) %>%
  mutate(Year = 2024) %>% 
  select(State, Year, Total_EE_Investment, Political_Score, Disasters, 
         `Total population`, Female, child, `Median age (years)`, 
         `Median household income (dollars)`, Gas, Electricity)

#Merge data
final_data <- bind_rows(df_2021, df_2022, df_2023, df_2024)
write.csv(final_data, file = "Data/Processed/final_data.csv")
```

# Rationale and Research Questions
As the threats of climate change continue to grow, green investment becomes increasingly important to mitigate its effects. Our team is interested in the factors driving federal investments in the energy and environmental sectors and whether it has changed over time. To analyze this question, we are using a dataset from the Federal Energy and Environment Investment Project to look at patterns in federal grant and loan spending from 2021 to 2024.  

We hypothesized that the frequency of natural disasters might draw more attention to investment into the energy and environmental sectors. Voter concern about climate change may also attract more federal investment. 

1. How does federal energy and environment investment change from 2021 to 2024?
2. What factors affect total federal energy and environment spending? 
3. Where is federal energy and environment spending being invested (i.e. where does the money go)?

\newpage

# Dataset Information
```{r}
summary(combined2021)
summary(census2021clean)
summary(Investment_state_2021)
summary(final_data)
```

Detail | Description
------------- | -------------
Data Source  | THE FEDERAL ENERGY AND ENVIRONMENT INVESTMENT PROJECT
Retrieved from  | https://www.pdx.edu/policy-consensus-center/federal-energy-environment-investment-project 
Variables Used  | Total population, Male, Female, Median age, Median household income, Gas, Electricity, All other fuels, No fuel used, Year, State, Child (age 0-17)
Date Range  | 2021-2024

Detail | Description
------------- | -------------
Data Source  | OpenFEMA Dataset: Disaster Declarations Summaries - v2
Retrieved from  | https://www.fema.gov/openfema-data-page/disaster-declarations-summaries-v2
Variables Used  | State, Disaster Number
Date Range  | 2021-2024

\newpage

# Exploratory Analysis 

```{r}
#create exploratory plot
cols <- c(1,3,6,7,9,10,15,18,24)

ggpairs(final_data, columns = cols, title = "",  
        axisLabels = "show",
        lower = list(continuous = wrap("points", size = 0.4)),
        upper = list(continuous = wrap("cor", size = 3)))+
  theme( #adjust the axis labels
    axis.text.x = element_text(size = 5),
    axis.text.y = element_text(size = 5),
    strip.text.x = element_text(size = 5),  
    strip.text.y = element_text(size = 5)
  )

 

```

```{r}
#visualize frequency of natural disasters 2021-2024
#summarize count 
disaster_count <-disasters %>% 
  filter(fyDeclared %in% c(2021, 2022, 2023, 2024)) %>% 
  arrange(state) %>% 
  group_by(state, fipsStateCode) %>% 
  summarise(Disasters = n())

disaster_count <- left_join(disaster_count, states_names, by = c("state" = "abbrev"))

#remove territories
disaster_count_clean <-disaster_count %>% 
    filter(fipsStateCode %in% c(1:56)) 
states_clean <-states %>% 
    filter(GEOID != 72) 


#join disaster to spatial data
states_sf_join <- states_clean %>%  
  left_join( 
    disaster_count_clean,  
    by = c("NAME" = "State")  
    ) 

st_crs(states_sf_join) #NAD83, espg 4269

#View with mapview - count of disasters
mapview(states_sf_join,  
        zcol = 'Disasters', # colors by disaster count
        col.regions = rev(brewer.pal(11, "RdBu"))) #high count in red 

#view with ggplot 
ggplot() +  
  geom_sf(data=states_sf_join,aes(fill = Disasters),alpha=0.5) +  
  scale_fill_gradient2(low="blue",high="red",midpoint = 0.5) + #high count in red 
  labs(fill='Disasters Frequency',
       title = "Disaster Frequency from 2021-2024")

#Texas, Florida, and Louisiana have the highest declared disasters from 2021-2024
```

\newpage

# Analysis

## Question 1: How does federal energy and environment investment change from 2021 to 2024?
```{r}
#time series analysis
```

## Question 2: What factors affect total federal energy and environment spending? 

This section investigated the drivers of federal energy and environmental investment across U.S. states in 2021, specifically examining the roles of environmental necessity (natural disasters) and political demand (public opinion).

```{r message=FALSE, warning=FALSE}
#Demographic Linear regression
#Some orginal data (eg. disaster) does not follow normal distribution, thus need to be log-transformed for linear regression.
log_combined2021<- combined2021%>%
  mutate(log_pop=log(`Total population`),
         log_disaster=log(Disasters),
         log_EE_investment=log(TotalFedAction_EnergyEnv_2021))

model_1 <- lm(data = log_combined2021, log_EE_investment ~ log_pop + Female + child + `Median age (years)` + `Median household income (dollars)`  + Gas + Electricity, na.action = na.exclude)
summary(model)
step(model_1)

model_2 <- lm(data = log_combined2021, TotalFedAction_EnergyEnv_2021 ~ log_pop + child + `Median age (years)`  + Electricity)
summary(model_2)
step(model_2)
#Among all demographic variables, Total population, electricity and median age population are significantly related to energy and environment investment


model_3 <- lm(data = log_combined2021, log_EE_investment ~  log_disaster, na.action = na.exclude)
summary(model_3) #frequency of disaster is related

model_4 <- lm(data = log_combined2021, log_EE_investment ~  `2021`, na.action = na.exclude)
summary(model_4) # Public awareness is related


plot1 <- ggplot(log_combined2021, aes(x=log_disaster, y = log_EE_investment))+
  geom_point()+
  geom_smooth(method = 'lm' )+
  ylab("Logarithm of Energy and Environment Investment")+
  xlab("Logarithm of the Frequency of Disasters across the US")
plot1

plot2 <- ggplot(log_combined2021, aes(x=`2021`, y = log_EE_investment))+
  geom_point()+
  geom_smooth(method = 'lm' )+
  ylab("Logarithm of Energy and Environment Investment")+
  xlab("")
plot2

plot3 <- ggplot(combined2021, aes(x=Disasters, y = TotalFedAction_EnergyEnv_2021))+
  geom_point()+
  geom_smooth(method = 'lm')
plot3 #plot2 is also an evidence showing why we log transformed the dataset


```

In model 1 & 2, preliminary demographic analysis was conducted to identify baseline socio-economic drivers. The results indicated that Total Population (log_pop) is the dominant predictor of federal investment (p < 0.001), while median household income showed no significant effect. This necessitates including population as a key control variable in subsequent hypothesis testing to avoid omitted variable bias.

In model 3, we tested whether disaster frequency alone predicts investment. The model showed a significant positive relationship (p < 0.001), initially supporting the hypothesis that funding is reactive. However, this simple model does not account for state size, raising the question of whether the funding is truly disaster-driven or merely proportional to population.

Similarly, Model 4 examined the relationship between public opinion on climate change and federal investment. The results indicated a significant positive correlation (p < 0.05), suggesting that states with higher climate awareness tend to attract more funding.

However, these models were limited by potential confounding factors (e.g., state size) and limited sample size (single year). To address this, we constructed a Final Integrated Model using integrated year data from 2021 to 2024.

## Question 3: Where is federal energy and environment spending being invested (i.e. where does the money go)?
```{r}
#This codes are for the merged (2021-2024) file analysis we can either choose analysis seperately or use model below!
final_data <- final_data %>%
  mutate(
    log_pop = log(`Total population`),
    log_disaster = log(Disasters + 1), 
    log_EE_investment = log(Total_EE_Investment)
  )

final_model <- lm(data = final_data, 
                  log_EE_investment ~ log_disaster + Political_Score + log_pop + as.factor(Year))

summary(final_model)
```

When analyzing the full panel data (2021-2024) and controlling for population size, the influence of disaster frequency (log_disaster) became statistically insignificant (p = 0.61). This contrasts with the preliminary analysis and suggests that the previously observed correlation between disasters and investment was largely driven by state size—larger states experience more disasters and receive more funding.

However, political pressure (Political_Score) remained a significant predictor (p < 0.05) even after controlling for population. This implies that states with a higher percentage of voters concerned about climate change actively attract more federal investment, independent of their size.

Furthermore, a temporal analysis reveals a significant surge in investment in 2024 (Coef = 0.237, p < 0.001) compared to the 2021 baseline.

\newpage

# Summary and Conclusions
1. Contrary to our initial hypothesis, natural disaster frequency is not a direct driver of federal investment when state size is taken into account.
2. Political demand for climate action effectively influences the distribution of federal resources.

The allocation of federal energy and environmental funds is best described as demographically scaled but politically responsive. While the baseline funding is determined by how many people live in a state, the variations above that baseline are significantly shaped by the political will of its voters, rather than the immediate frequency of natural disasters.


\newpage

# References
<add references here if relevant, otherwise delete this section> 
