---
output: html_document
editor_options: 
  chunk_output_type: console
---
\newpage
\tableofcontents 
\newpage
\listoftables 
\newpage
\listoffigures 
\newpage

```{r setup, include=FALSE}
# Set your working directory
# Load packages
getwd()
library(tidyverse)
library(sf)
library(ggplot2)
library(dplyr)
library(mapview)
library(leaflet)
library("GGally")
library(readxl)

# Set theme
mytheme <- theme_classic(base_size = 14) +
  theme(axis.text = element_text(color = "black"), 
        legend.position = "top")
theme_set(mytheme)

```

```{r}
# Read in census data
census_processed<- read.csv("Data/census_processed(2020-2024).csv", stringsAsFactors = T)

census_processed <- census_processed %>% 
  mutate(MEAN = rowMeans(across(X2020:X2024)))
substring(census_processed$State, 2, nchar(census_processed))
colnames(census_processed) <- c("State","2020","2021","2022","2023","2024")

#Federal Energy and Environment investment data by county 
Energy_investment<- read.csv("Data/Energy_investment_by_county.csv", stringsAsFactors = T)

#Federal Energy and Environment investment data by state 
Investment_state <- read.csv("./Data/Raw/Federal energy and environment spending by STATE (2010-2025).csv", stringsAsFactors = T)

#Natural disaster data
disasters <- read.csv("./Data/DisasterDeclarationsSummaries.csv", stringsAsFactors = T)

#spatial data
states <- st_read("./Data/Spatial/cb_2018_us_state_20m.shp", stringsAsFactors = T)


census_2021 <- read.csv("ACSSPP1Y2021.S0201-2025-11-29T154845.csv", stringsAsFactors = T)
census_2022 <- read.csv("ACSSPP1Y2022.S0201-2025-11-29T154817.csv", stringsAsFactors = T)
census_2023 <- read.csv("ACSSPP1Y2023.S0201-2025-11-29T154759.csv", stringsAsFactors = T)
census_2024 <- read.csv("ACSSPP1Y2024.S0201-2025-11-29T154742.csv", stringsAsFactors = T)
```



```{r}
#Data Wrangling (for 2021)
##Clean the data

#census_2021.columns = census_2021.columns.str.replace(' ','')
census2021 <- as.data.frame(t(census_2021))
colnames(census2021) <-c(census2021[1,])

census2021clean <- census2021[-1,]
#census2021clean <- census2021[-c(seq(2,length(census2021),2)),]

colnames(census2021clean) <- str_trim(colnames(census2021clean))

rownames(census2021clean) <- str_sub(rownames(census2021clean), end =-29)
#rownames(census2021clean) <- gsub("."," ", rownames(census2021clean))

census2021clean <- census2021clean %>% 
  select("Total population", "Male", "Female", "Under 5 years", "5 to 17 years", "Median age (years)", "Median household income (dollars)","Gas", "Electricity","All other fuels","No fuel used") %>% 
  mutate(year = 2021, State = rownames(census2021clean))
  #mutate(State = sub("."," ", State)) # Remove the dots in State names
census2021clean$State <- gsub( "\\."," ",census2021clean$State)

census2021clean$`Total population` <- as.numeric(str_trim(census2021clean$`Total population`))
census2021clean$`Male` <- as.numeric(str_trim(census2021clean$`Male`))
census2021clean$Female <- as.numeric(str_trim(census2021clean$`Female`))
census2021clean$`Under 5 years` <- as.numeric(str_trim(census2021clean$`Under 5 years`))
census2021clean$`5 to 17 years` <- as.numeric(str_trim(census2021clean$`5 to 17 years`))
census2021clean$`Median age (years)` <- as.numeric(str_trim(census2021clean$`Median age (years)`))
census2021clean$`Median household income (dollars)` <- as.numeric(str_trim(census2021clean$`Median household income (dollars)`))

census2021clean$Gas <- as.numeric(str_trim(census2021clean$Gas))
census2021clean$Electricity <- as.numeric(str_trim(census2021clean$Electricity))
census2021clean$`All other fuels` <- as.numeric(str_trim(census2021clean$`All other fuels`))
census2021clean$`No fuel used` <- as.numeric(str_trim(census2021clean$`No fuel used`))


#add column for children ages 0-17 

census2021clean <- census2021clean%>%
  mutate(child=`Under 5 years` + `5 to 17 years`)


## Energy dataset (State Level)
Investment <- Investment_state %>%  #data for time series
  select(
    "TotalFedAction_EnergyEnv_2010","TotalFedAction_2010",
    "TotalFedAction_EnergyEnv_2011","TotalFedAction_2011",
    "TotalFedAction_EnergyEnv_2012","TotalFedAction_2012",
    "TotalFedAction_EnergyEnv_2013","TotalFedAction_2013",
    "TotalFedAction_EnergyEnv_2014","TotalFedAction_2014",
    "TotalFedAction_EnergyEnv_2015","TotalFedAction_2015",
    "TotalFedAction_EnergyEnv_2016","TotalFedAction_2016",
    "TotalFedAction_EnergyEnv_2017","TotalFedAction_2017",
    "TotalFedAction_EnergyEnv_2018","TotalFedAction_2018",
    "TotalFedAction_EnergyEnv_2019","TotalFedAction_2019",
    "TotalFedAction_EnergyEnv_2020","TotalFedAction_2020",
    "TotalFedAction_EnergyEnv_2021","TotalFedAction_2021",
    "TotalFedAction_EnergyEnv_2022","TotalFedAction_2022",
    "TotalFedAction_EnergyEnv_2023","TotalFedAction_2023",
    "TotalFedAction_EnergyEnv_2024","TotalFedAction_2024") %>% 
  mutate(across(contains("FedAction"),
                ~ as.numeric(str_replace_all(.x, "[\\$,]", "")))) 
  
  
Investment_state_2021 <- Investment_state %>% 
  select("STATEFP","stname", "stusps","TotalFedAction_EnergyEnv_2021","TotalFedAction_2021")
Investment_state_2021$TotalFedAction_EnergyEnv_2021 <- 
gsub("\\$ ", "", Investment_state_2021$TotalFedAction_EnergyEnv_2021)
Investment_state_2021$TotalFedAction_2021 <- 
gsub("\\$ ", "", Investment_state_2021$TotalFedAction_2021)
Investment_state_2021 <- Investment_state_2021 %>% 
  mutate(TotalFedAction_EnergyEnv_2021 =
           as.numeric(
             gsub(",", "", trimws(TotalFedAction_EnergyEnv_2021)))
         ) %>% 
  mutate(TotalFedAction_2021 =
           as.numeric(
             gsub(",", "", trimws(TotalFedAction_2021)))
         ) %>% 
  mutate(Ratio = TotalFedAction_EnergyEnv_2021/TotalFedAction_2021,
         Year = 2021)


#Disaster
states_names <- data.frame(abbrev = states$STUSPS,
                           state = states$NAME) 
#The disaster data set has only abbreviation for states, so full names need to be joined from other dataset.
disaster2021 <- disasters %>% 
  select(c(state:incidentType), fipsStateCode,fipsCountyCode) %>% 
  filter(fyDeclared %in% c(2021)) %>% 
  arrange(state) %>% 
  group_by(state) %>% 
  summarise(Disasters = n())
disaster2021 <- left_join(disaster2021, states_names, by = c("state" = "abbrev"))


#Public awareness
Regulate <- read_excel("Data/Regulate_percentage(2021~2024).xlsx")
Regulate2021 <- Regulate %>% 
  select("GeoName","2021")


#Combining all the variables (Demographic, disaster and public awareness)
combined2021 <- left_join(census2021clean, Investment_state_2021, by = c("State" = "stname"))

combined2021 <- left_join(combined2021, disaster2021, by = c("State" = "state.y"))

combined2021 <- left_join(combined2021, Regulate2021, by = c("State" = "GeoName"))


write.csv(combined2021, file = "./Data/Processed/combined2021.csv")
```

# Rationale and Research Questions

1. How does federal energy and environment investment change from 2021 to 2024?
2. What factors affect total federal energy and environment spending? 
3. Where is federal energy and environment spending being invested (i.e. where does the money go)?

\newpage

# Dataset Information
```{r}
summary(combined2021)
summary(census2021clean)
summary(Investment_state_2021)
```

Detail | Description
------------- | -------------
Data Source  | THE FEDERAL ENERGY AND ENVIRONMENT INVESTMENT PROJECT
Retrieved from  | https://www.pdx.edu/policy-consensus-center/federal-energy-environment-investment-project 
Variables Used  | Total population, Male, Female, Median age, Median household income, Gas, Electricity, All other fuels, No fuel used, Year, State, Child (age 0-17)
Date Range  | 2021-2024

Detail | Description
------------- | -------------
Data Source  | OpenFEMA Dataset: Disaster Declarations Summaries - v2
Retrieved from  | https://www.fema.gov/openfema-data-page/disaster-declarations-summaries-v2
Variables Used  | State, Disaster Number
Date Range  | 2021-2024

\newpage

# Exploratory Analysis 

```{r}
#create exploratory plot
cols <- c(1,3,6,7,9,10,15,18,20,24)

ggpairs(combined2021, columns = cols, title = "",  
        axisLabels = "show",
        lower = list(continuous = wrap("points", size = 0.4)),
        upper = list(continuous = wrap("cor", size = 3)))+
  theme( #adjust the axis labels
    axis.text.x = element_text(size = 5),
    axis.text.y = element_text(size = 5),
    strip.text.x = element_text(size = 5),  
    strip.text.y = element_text(size = 5)
  )

 #Some orginal data (eg. disaster) does not follow normal distribution, thus need to be log-transformed for linear regression.

```


\newpage

# Analysis

## Question 1: How does federal energy and environment investment change from 2021 to 2024?
```{r}
#time series analysis
```

## Question 2: What factors affect total federal energy and environment spending? 

```{r message=FALSE, warning=FALSE}
#Demographic regression
log_combined2021<- combined2021%>%
  mutate(log_pop=log(`Total population`),
         log_disaster=log(Disasters),
         log_EE_investment=log(TotalFedAction_EnergyEnv_2021))

model <- lm(data = log_combined2021, log_EE_investment ~ log_pop + Female + child + `Median age (years)` + `Median household income (dollars)`  + Gas + Electricity, na.action = na.exclude)
summary(model)
step(model)

model_2 <- lm(data = log_combined2021, TotalFedAction_EnergyEnv_2021 ~ log_pop + child + `Median age (years)`  + Electricity)
summary(model_2)
step(model_2)
#Among all demographic variables, Total population, electricity and median age population are significantly related to energy and environment investment


model_3 <- lm(data = log_combined2021, log_EE_investment ~  log_disaster, na.action = na.exclude)
summary(model_3) #frequency of disaster is related

model_4 <- lm(data = log_combined2021, log_EE_investment ~  `2021`, na.action = na.exclude)
summary(model_4) # Public awareness is related


plot1 <- ggplot(log_combined2021, aes(x=log_disaster, y = log_EE_investment))+
  geom_point()+
  geom_smooth(method = 'lm' )+
  ylab("Logarithm of Energy and Environment Investment")+
  xlab("Logarithm of the Frequency of Disasters across the US")
plot1

plot2 <- ggplot(log_combined2021, aes(x=`2021`, y = log_EE_investment))+
  geom_point()+
  geom_smooth(method = 'lm' )+
  ylab("Logarithm of Energy and Environment Investment")+
  xlab("")
plot2

plot3 <- ggplot(combined2021, aes(x=Disasters, y = TotalFedAction_EnergyEnv_2021))+
  geom_point()+
  geom_smooth(method = 'lm')
plot3 #plot2 is also an evidence showing why we log transformed the dataset


```

## Question 3: Where is federal energy and environment spending being invested (i.e. where does the money go)?
```{r}

```


\newpage

# Summary and Conclusions


\newpage

# References
<add references here if relevant, otherwise delete this section> 
